Project Proposal

### 2024-11-16 proposal_EDA/deGuzman_IDS702Projectproposal_template.qmd
updating proposal draft

---
title: "Project Proposal"
author: "Sizhe Chen, Naikyah Dhariwala, Peter de Guzman, Jennifer Li"
subtitle: "Due November 17 at 11:59pm"
format: pdf
---

@@

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse)
```

@@

@@

```{r data-1, warning=FALSE}
rm(list=ls())
load("/Users/pdeguz01/Documents/git/Data/2020 CMPS/DS0001/39096-0001-Data.rda")
data <- da39096.0001
data %>% slice_head(n = 100) %>% glimpse()
```

@@

```{r plots-1}
wtd_tableQ29 <- data %>% count(Q29, wt=WEIGHT)
wtd_tableQ29 <- wtd_tableQ29 %>% mutate('pct' = (n/sum(n, na.rm=TRUE)))

@@

contingency_table <- table(data$Q29, data$Q271)
contingency_df <- as.data.frame(contingency_table)
contingency_df <- contingency_df %>% mutate(Share = (Freq/sum(Freq)))

ggplot(contingency_df, aes(x = Var1, y = Var2, fill = Share)) +
  geom_tile() +
  labs(title="Heatmap of Interest in Politics and Racial Identity", x = "Interest in Politics", y = "Importance of Racial Identity") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_continuous(low = "white", high = "blue", labels = percent_format(scale = 100)) 
```

>>>>

---
title: "Project Proposal"
author: "Sizhe Chen, Naikyah Dhariwala, Peter de Guzman, Jennifer Li"
subtitle: "Due November 17 at 11:59pm"
format: pdf
header-includes:
      - \usepackage{float}
      - \floatplacement{table}{H}
---

@@

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse)
library(ggplot2)
library(formattable)
library(scales)
```

@@

The main predictor variable is "Q271" which asks “How important is being [RACIAL GROUP] to your identity?”. The response options are a five-point scale from "Not at all important" to "Extremely important". Other potential predictors include survey questions on respondent's self-perception of race and discrimination, and their own attitudes towards other racial groups. In addition, there are multiple variables that could be used as demographic controls such as age, gender, education, income, and party identification. The variables not shown in the output include responses to the rest of the 800 survey questions, covering topics such as policy attitudes, civic engagement, and political participation.

@@

```{r data-1, warning=FALSE, echo=FALSE}
rm(list=ls())
load("/Users/pdeguz01/Documents/git/Data/2020 CMPS/DS0001/39096-0001-Data.rda")
data <- da39096.0001
reordered_data <- data %>% relocate(Q29, Q271, WEIGHT, 
                                    AGE1_HID, AGE2_HID, .after = DATE)
reordered_data %>% slice_head(n = 50) %>% str()
```

@@

```{r plots-1, echo=FALSE}
wtd_tableQ29 <- data %>% count(Q29, wt=WEIGHT)
wtd_tableQ29 <- wtd_tableQ29 %>% mutate('pct' = (n/sum(n, na.rm=TRUE)))

@@

data <- data %>% mutate(Q29_shortened = case_when(Q29 == "(1) Very interested in politics" ~ "Very interested",
                                                   Q29 == "(2) Somewhat interested" ~ "Somewhat interested",
                                                   Q29 == "(3) Not that interested in politics" ~ "Not that interested",
                                                   Q29 == "(4) Not at all interested in politics" ~ "Not at all interested",
                                                   .default = "Other"))
contingency_table <- table(data$Q29_shortened, data$Q271)
contingency_df <- as.data.frame(contingency_table)
contingency_df <- contingency_df %>% mutate(Share = (Freq/sum(Freq)))

ggplot(contingency_df, aes(x = Var1, y = Var2, fill = Share)) +
  geom_tile() +
  labs(title="Interest in Politics and Racial Identity", x = "Interest in Politics", y = "Importance of Racial Identity") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 0.4, vjust = 0.5)) +
  scale_fill_continuous(low = "white", high = "blue", labels = percent_format(scale = 100)) 
```
